\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-08-01 paper/ResponseToReviewers.tex

Reviewer 1 was concerned with some decisions that we had made in our computational modeling and statistical analyses and the interpretation of the results. In response to the concerns about our statistical analyses, we have completely re-done the analyses of the reading times and ERP data and added analyses of the previous data. Additionally, we have added one computational model and use k-fold cross-validation for model comparisons. We believe that these new analyses strengthen our initial interpretation that our study shows semantic interference, but no syntactic interference and no interaction.

Reviewer 2 primarily raised concerns regarding our experimental design. In response to the concerns about our experimental design, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textit{Journal of Memory and Language}. We now stress in the paper that we are building on previously published work by directly comparing the statistical inferences, using Bayes factors, from the present and previous studies that use exactly the same design.

Reviewer 3 suggested a more in-depth discussion of the mechanisms of cue-based retrieval and had some concerns about our methodology. We have added a more detailed discussion of the memory mechanisms which are assumed to underlie sentence processing in the present design. We have added a thorough discussion of all concerns which were raised regarding our methodology.

Below, we separately address each comment from the reviewers and yourself. All changes have been highlighted in the revised manuscript. The actionable parts of the comments are highlighted in bold.

@@

The design of the current studies repeats a design used in previous studies in German and other languages. The more novel aspects of the current studies are the use of the ERP measure, and the inclusion of a very large number of participants, allowing for more precise estimates of the syntactic and semantic interference effects. The quantitative analyses are characteristically thorough and careful.

I would support publication in JML. But I think that the paper would be stronger if there was fuller engagement with the range of possible reasons for the observed effects. This covers (i) mechanisms that could capture the lack of a syntactic interference effect, and (ii) reasons for differences between the current study and others where a syntactic interference effect was found.

The authors attribute the lack of a syntactic interference effect to the use of a syntactic feature like ``subject of the same clause", which has the effect that a subject of a different clause does not count as similar to the retrieval cue. This gives a special status to the correct subject via the choice of retrieval cues. \textbf{There may be other ways of giving special status to the correct subject, and thereby capturing the lack of syntactic interference. For example, if the subject of the current clause is encoded in a special status, e.g., with boosted activation, or with its needs prioritized, then this could have a similar effect.}''
\end{quote}

\subsubsection*{Response to R3.1}
It is possible that the subject of the current clause receives an activation boost but the mechanism for that is not trivial. Let's assume that the subject is encoded with boosted activation compared to for example an object. As the sentence progresses the subject's activation decays (we could even assume that it decays more slowly than other memory representations due to its special status) and other nouns are processed / encoded. At one point within a long sentence, the subject's activation will be lower than the activation of other nouns in memory. If we assume that a long sentence has multiple embedded sentences, we would need to assume that either the matrix sentence subject activation is downgraded when an embedded sentence is entered or that the embedded sentence subject gets an even higher activation boost than the matrix sentence subject. Then when the embedded sentence is finished and we return back to the matrix sentence, the matrix sentence subject would need to receive another activation boost to return to its special status. Essentially, this proposal condenses down to a clause tracking mechanism with a modulation of activation levels. We think it is more parsimonious to implement the clause tracking in the memory encoding and retrieval operations than the activation levels.

@@

\subsection*{Comment R3.2}
\begin{quote}
``The use of a feature like ``subject of the same clause" calls for some further discussion, since, taken literally, this is not a likely feature of a subject phrase in memory. Features like ``animate" or ``plural" are inherent properties of an item in memory, but ``subject of the same clause" is not. \textbf{I imagine that what the authors have in mind is that comprehenders assign a clause index to each individual piece of structure as they parse a sentence. Then, when they encounter a verb, they could identify the current clause index, e.g., ``3", and then embed that in a retrieval cue, e.g., ``subject of clause 3". This is a fine way to introduce a limited amount of relational information into a structural encoding. But note that it means that a good deal of work is now done by the mechanism that continually tracks clause indices and keeps track of the index of the current clause. So, it amounts to a mechanism that does more than simply elaborate the choice of retrieval cues at the verb.}''
\end{quote}

\subsubsection*{Response to R3.2}
The Reviewer is correct, our proposal that hierarchical syntactic structure is used during retrieval relies on the general assumption that hierarchical syntactic structure is being tracked. We think that this can easily be represented in encodings and then be used during retrieval. We have added a more explicit description of this underlying mechanism to page \pageref{clause_tracker}:

\Paste{clause_tracker}

@@

\subsubsection*{Response to R3.4}
As described in our Response to R1.4, we have added a detailed comparison of our results to the previous results. The main difference is that our data showed no evidence for syntactic interference in the pre-critical region, while the previous studies did show support for it. We have created a new subsection in our General Discussion providing an extended discussion of the potential reasons for the differences between our and previous results. This discussion did not lead to a definite answer. We provide the relevant new text passages in our responses below (Responses to R3.7 and R3.8).

We think that we have addressed the concerns regarding high- vs.\ lower-powered studies and contradictory results between studies in our Response to R1.5. We do not claim that our results are correct and previous claims have been false. Progress in science is achieved by carrying out adequately powered studies in the future that can
distinguish between theoretical claims. The replication crisis \citep{openscience2015_reproducibility} has shown that contradiction between results even when using the same design is not an anomaly; if anything, it's quite the opposite.

\subsection*{Comment R3.5}
\begin{quote}
``This may be an over-simplistic approach to the findings. First, \textbf{could there have been an impact of the fact that the study included an unusually large number of experimental items with identical structures, leading to practice effects?}''
\end{quote}

\subsubsection*{Response to R3.5}
To account for practice effects, we have added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data. Theoretically, participants could either speed up or slow down (or do not change their reading speed) over the course of the experiment. Therefore, a symmetrical prior on the slope of trial id (effect could be positive, negative or zero) is the most reasonable. We decided to use a Normal(0, 0.1) prior on the trial id slope in all models, a priori assuming effects up to $\pm$ 81\,ms. For the first trial in the experiment to the last trial, participants actually sped up drastically: Reading times in the regions of interest decreased by on average around 250\,ms.

It is not possible to use symmetrical priors (for trial id) and directional priors (which we had used originally for the effects of interest to reflect longer reading times in high interference conditions) within the same brms model \citep{brms}. So in the revised analyses, we now use symmetrical priors on all slopes which are a priori agnostic regarding the sign of the effect. To ensure consistency in our analyses and comparability between the analyses, we changed all our analyses (SPR and ERP data) to use symmetrical priors. These revised analyses have produced the same result pattern as our previous analyses (no syntactic interference or interaction, but a semantic interference effect).

### 2024-08-09 paper/main.tex
Updated via the Open Science Framework

\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?
\newline
Evidence from self-paced reading and event-related potentials using German}
\authorsnames{Pia Schoknecht, Shravan Vasishth}

\shorttitle{Syntactic and semantic interference effects}
\affiliation{
Department of Linguistics, University of Potsdam}

\authornote{This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287.
We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition. We are grateful to Dario Paape for helpful discussions, and Himanshu Yadav for assistance with the computational modeling. 
Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
}

@@

Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject. \Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary}.}}}\label{franck_refs}

Semantic interference is manipulated via the animacy of the distractor. During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference. \Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained. In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}. While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant. The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval. A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
}}\label{semantic_features_encoding}

@@

For these reasons, we added two adverbs before the critical verb. Including these intervening regions had a further advantage: 
It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region. The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power. 
Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region. This way, the critical region should be less affected by the clause-boundary slowdown. This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.

Thus, our focus in the \textcolor{blue}{SPR experiment} is on the processing difficulty observed at the critical verb and the surrounding regions.

@@

\textcolor{blue}{Alternately, interference -- especially syntactic interference -- might modulate the P600. The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset. It is associated with syntactic processing, e.g., morpho-syntactic errors and reanalysis \citep[see, e.g.,][]{osterhout&holcomb_1992, kaan_etal_2000_P600}. High interference could increase difficulty during syntactic integration of subject and verb in our materials. Potential misretrievals of the distractor might lead to reanalysis processes which could manifest as an increased P600 amplitude compared to low interference conditions. To anticipate our results here, visual inspection of the ERPs in Figure \ref{fig:erp_syn_ERP} does not indicate syntactic interference effects in the P600 time window.}} \label{why_n400}

@@

Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. \textcolor{blue}{In addition to the predictors of interest, we included the trial id to account for potential effects of fatigue or adaptation to the task. The trial id of the 100 trials per session was recoded to span from 0 to 1. This was done to bring all fixed effects to a comparable scale which decreases the run time of the models. The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items).}

@@

\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method. A Bayesian hypothesis test for which the posterior density is devided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$. \textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account. This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}. Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect. This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.

@@

Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters. \Copy{SPR_results}{\textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions. In the spill-over region, it included only positive values ([1, 15] ms). In contrast, for semantic interference, the 95 \% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms). The posteriors for the interaction were mostly positive, but the 95 \% credible intervals included zero in all regions. The interaction was the  largest in the critical region ([-3, 31 ms]).}

@@

\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last, i.e., 100th, trial, reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]). This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment. While this adaption is interesting, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}

@@

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5). There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was weak evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

@@

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{Reading times of the pre-critical word and critical verb showed evidence against syntactic interference. In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms. We will discuss the almost complete absence of syntactic interference in our data in the general discussion. The reading time results at the critical verb were inconclusive for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region. This means that  semantic interference occurred even before the retrieval event could begin. Reading times at the critical verb showed very weak support for a small effect of a syntactic $\times$ semantic interference interaction. There was evidence against the interaction under broader priors and in the surrounding regions.}\Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.

@@

\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience). }

@@

\textcolor{blue}{Forty-four native speakers of German (mean age: 26 years old, age range: 18 - 35 years, 21 female, 23 male), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible). The mean ratings per item ranged from 3.6 to 6.6. The experiment included 40 implausible filler sentences which were created by exchanging nouns between the fillers used in the SPR experiment (e.g., Der Sohn des Kaisers, der den Feldzug gewonnen hatte, plante äußerst strategisch, weil er die Gießkanne an sich reißen wollte., `The son of the emperor who had won the campaign planned extremely strategically because he wanted to take the watering can by force.').  The mean rating of the implausible fillers ranged from 1.1 to 3.2. Participants spent approximately 37 minutes to complete the task and received \pounds 6.50 as compensation. The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}.}

@@

\textcolor{blue}{Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible. To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. For the slopes, we used a Normal(0,0.5) prior; the priors for all other parameters were the same as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from this model. Although an increase in plausibility lead to faster reading times (CrI [-30, -6] ms), there was still an effect of semantic interference (CrI [2, 24] ms). The Bayes factor provided weak evidence for semantic interference in this analysis (BF$_{10}$ = 2.4). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}\label{plausib_anim_inan}

@@

Figure \ref{fig:erp_all} shows the 
grand average ERPs elicited by the critical verb for all conditions. \textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$. }\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600. The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals include zero for both the N400 and the P600. The 95\% credible interval of the interaction was centered around zero for the N400. For the P600, it included mostly negative values but crossed zero.}}\label{ERP_results1}

@@

\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$). In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference. Both the N400 and P600, were more negative under high vs. low semantic interference. In contrast, there was no evidence or even evidence against effects of syntactic interference and the interaction on the N400 and P600. In sum, the ERP results showed clear evidence for semantic interference and no evidence or evidence against syntactic interference and the interaction of syntactic and semantic interference. Consequently, we conclude that the ERP data showed no syntactic interference and no interaction effect.}}\label{ERP_discussion}

>>>>

\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?
\newline
Evidence from self-paced reading and event-related potentials using German}

\authorsnames[1,{2},1]{Pia Schoknecht, Himanshu Yadav, Shravan Vasishth}
\authorsaffiliations{{Department of Linguistics, University of Potsdam},
{Department of Cognitive Science, Indian Institute of Technology Kanpur}}

\shorttitle{Syntactic and semantic interference effects}

\authornote{This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287.
We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition. We are grateful to Dario Paape for helpful discussions.
Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
}

@@

Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject. \Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary, franck2020hierarchical}.}}}\label{franck_refs}

Semantic interference is manipulated via the animacy of the distractor. During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference. \Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained. In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}. While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant. The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval (for work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \cite{vandyke_mcelree06, cunnings_sturt_2018retrieval}). A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
}}\label{semantic_features_encoding}

@@

For these reasons, we added two adverbs before the critical verb. Including these intervening regions had a further advantage: 
It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region. The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power. 
Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region. This way, the critical region should be less affected by the clause-boundary slowdown. This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region. Thus, our focus in the \textcolor{blue}{SPR experiment} is on the processing difficulty observed at the critical verb and the surrounding regions.

@@

\textcolor{blue}{Alternately, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600. The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset. It is associated with syntactic processing, e.g., morpho-syntactic errors and reanalysis \citep[see, e.g.,][]{osterhout&holcomb_1992, kaan_etal_2000_P600}. High interference could increase difficulty during syntactic integration of subject and verb in our materials. Potential misretrievals of the distractor might lead to reanalysis which could manifest as an increased P600 amplitude in high compared to low interference conditions.}} \label{why_n400}

@@

Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. \textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task. The trial id of the 100 trials per session was recoded to span from 0 to 1. This was done to bring all fixed effects to a comparable scale which decreases the run time of the models. The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items).}

@@

\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method. A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$. \textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account. This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}. Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect. This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.

@@

Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters. \Copy{SPR_results}{\textcolor{blue}{The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions. In the spill-over region, it included only positive values ([1, 15] ms). In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms). The posteriors for the interaction were mostly positive, but the 95\% credible intervals included zero in all regions. The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}

@@

\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]). This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment. While this adaption is interesting in itself, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}

@@

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

@@

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{Reading times of the pre-critical word and critical verb showed evidence against syntactic interference. In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms. We will discuss the almost complete absence of syntactic interference in our data in the general discussion. The reading time results at the critical verb were inconclusive for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region. This means that semantic interference occurred even before the retrieval event could begin. Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference. There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.

@@

\label{plausib_anim_inan}\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience). }

@@

\textcolor{blue}{Forty-four native speakers of German (mean age: 26 years old, age range: 18 - 35 years, 21 female, 23 male), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible). The mean ratings per item ranged from 3.6 to 6.6. The experiment included 40 implausible filler sentences which were created by exchanging nouns between the fillers used in the SPR experiment (e.g., Der Sohn des Kaisers, der den Feldzug gewonnen hatte, plante äußerst strategisch, weil er die Gießkanne an sich reißen wollte., `The son of the emperor who had won the campaign planned extremely strategically because he wanted to take the watering can by force.').  The mean rating of the implausible fillers ranged from 1.1 to 3.2. Participants spent approximately 37 minutes to complete the task and received \pounds 6.50 as compensation. The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}. Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.}

@@

\textcolor{blue}{To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}

@@

Figure \ref{fig:erp_all} shows the 
grand average ERPs elicited by the critical verb for all conditions. \textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$. }\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600. The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals included zero for both the N400 and the P600. The 95\% credible interval of the interaction was centered around zero for the N400. For the P600, it included mostly negative values but crossed zero.}}\label{ERP_results1}

@@

\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$). In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference. Both the N400 and P600 were more negative under high vs. low semantic interference. In contrast, there was only very weak evidence for very small effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows. In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference. Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}}\label{ERP_discussion}
